Improving the Performance of Distributed Data Parallelism \\Over Low Communication Bandwidths

### 2023-02-20 paper/paper.tex
updates

Modern models can contain billions of trainable parameters, which translates to billions of gradients. If we represent all of those gradient as 256 different integers we loose a lot of accuracy and make it alot harder for the model to converge, so what we do is we split gradients into chunks of around 2000 or 4000 items(depending on the size of the model) and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk. There are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantization step a lot faster. If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins it to be wasted because of single outlier. By splitting up the data into chunks only a one chunk would be effected by this outliers.

@@

In order to perform gradient sparsification we have to decide on a threshold to determine whether or not a gradient is a zero gradient or not, if the gradient is less than the threshold then it it deemed a zero gradient. In our implementation of gradient sparsification we accumulate zero gradients locally and sum them with gradients from the next iteration.

>>>>

Modern models can contain billions of trainable parameters, which translates to billions of gradients. If we represent all of those gradients as 256 different integers we loose a lot of accuracy and make it alot harder for the model to converge, so what we do is we split gradients into chunks of around 2000 or 4000 items(depending on the size of the model) and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk. There are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantization step a lot faster. If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins which means it would be wasted because of single outlier. By splitting up the data into chunks only a one chunk would be effected by this outliers.

@@

In order to perform gradient sparsification we have to decide on a threshold to determine whether or not a gradient is a zero gradient. If the gradient is less than the threshold then it is deemed a zero gradient. In our implementation of gradient sparsification we accumulate zero gradients locally and sum them with gradients from the next iteration.
